---
title: "Analysis of US Storm Data for Human and Economic Impact"
author: "Manu Perez"
date: "February 15, 2016"
output: html_document
---

##### Which relation exists between severe weather events and humans? What is the economic impact?

##### This questions and others can be answered by statistic. And, may be in the future, we would can prevent the outcomes to the extent possible.

##### In this paper, we will try to answere there questions. For this, we will use the US Storm data from the NOAA Storm Database.


### Data Processing
##### First, we load the data. You can download the data from https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2

```{r}
cache=TRUE
path <- "D:/Cursos/Coursera/Curso Data Science Specialization 5. Reproducible Research/week 4/"
stormdata <- read.csv(paste(path,"repdata-data-StormData.csv", sep=""))
```

##### Because in the earlier years of the database there are generally fewer events recorded, we will work only with data of the last 20 years.

```{r}
stormdata$YEAR <- as.numeric(format(as.Date(stormdata$BGN_DATE, '%m/%d/%Y'), '%Y'))
stormdata20 <- stormdata[stormdata$YEAR >= 1981,]

```

##### We also clean data that did not have any property damages or crop damage and no fatalities or injuries.

```{r}
stormdata20 <- stormdata20[(stormdata20$PROPDMG > 0 | stormdata20$CROPDMG > 0 | stormdata20$FATALITIES > 0 | stormdata20$INJURIES > 0),]

```

##### We will aggregate the data by summing up the various Values by Event Type

```{r}
stormdataAggregated <- aggregate(cbind(stormdata20$PROPDMG, stormdata20$CROPDMG, stormdata20$FATALITIE, stormdata20$INJURIES) ~ stormdata20$EVTYPE, stormdata20, sum)

```

### Results

##### We only visualize the first 5 results.
##### With respect to property damage, TORNADO is the most dangerous weather event.

```{r}
stormdataPD <- stormdataAggregated[order(-stormdataAggregated[,"V1"]),]
stormdataPD <- stormdataPD[1:5,]
stormdataPD
```

##### Plot the data
```{r}
legPD <- as.vector(stormdataPD[,1])
barplot(c(as.numeric(unlist(stormdataPD["V1"]))), legend.text=legPD, main="Events with Highest Property Damage", col=heat.colors(5))

```

##### With respect to crop damage, HAIL is the most dangerous weather event.

```{r}
stormdataCD <- stormdataAggregated[order(-stormdataAggregated[,"V2"]),]
stormdataCD <- stormdataCD[1:5,]
stormdataCD
```
##### Plot the data
```{r}
legCD <- as.vector(stormdataCD[,1])
barplot(c(as.numeric(unlist(stormdataCD["V2"]))), legend.text=legCD, main="Events with Highest Crop Damage", col=heat.colors(5))
```

##### With respect to population impact, first, we combine fatalities and injuries.

```{r}
stormdataAggregated$POPIMP <- stormdataAggregated[,"V3"] + stormdataAggregated[,"V4"]
stormdataPI <- stormdataAggregated[order(-stormdataAggregated$POPIMP),]
stormdataPI <- stormdataPI[1:5,]
stormdataPI
```

##### Then, we can see that TORNADO is the most important population impact.

##### Plot the data
```{r}
legPI <- as.vector(stormdataPI[,1])
barplot(c(as.numeric(unlist(stormdataPI$POPIMP))), legend.text=legPI, main="Events with Highest Population Impact", col=heat.colors(5))
```
